---
title: "02 Spatial Interpolation on Feral Pig Abundance"
author: "Shuhan Song"
date: "11/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

### Attach packages

```{r}
library(raster)
library(tidyverse)
library(here)
library(lubridate)

# spatial analysis packages
library(rgdal)
library(tmap)
library(sf)
library(spatstat) 
library(maptools)
library(gstat)
library(sp)
library(rgeos)
```

### Read in spatial data

```{r}
# load preserve boundary data
jldp_boundary <- read_sf(dsn = here::here("raw_data", "jldp_boundary"), layer = "jldp_boundary")

# ---------------------------- IMPORTANT-------------------------------------------
# load abundance data
# september check abundance
september_abundance <- read_sf(dsn = here::here("cleaned_data", "PigAbundancePerSite_Sep"),
                      layer = "PigAbundancePerSite_Sep") %>% 
  dplyr::select(camer_d, wght_s_) %>% 
  rename(wght_abund = wght_s_) %>% 
  mutate(month_check = "September Check")

st_crs(september_abundance) = st_crs(jldp_boundary)

# april check abundance
april_abundance <- read_sf(dsn = here::here("cleaned_data", "PigAbundancePerSite_April"),
                      layer = "PigAbundancePerSite_April") %>% 
  dplyr::select(camer_d, wght_s_) %>% 
  rename(wght_abund = wght_s_)%>% 
  mutate(month_check = "April Check")

st_crs(april_abundance) = st_crs(jldp_boundary)
# ---------------------------- IMPORTANT-------------------------------------------
```

```{r}
# ---------------------------- IMPORTANT-------------------------------------------
# things to change:

# month of check (i.e April or September)
# input of abundance data (i.e. april_abundance or september_abundance)
# also rename attributes as needed
month_check = "April"
first_day = min(april_abundance$date)
last_day = max(april_abundance$date)

pig_abundance <- april_abundance %>% 
  rename(statn_d = camer_d, # rename attribute of camera id
         month_check = month_check, # rename the attribute of april check or september check
         nmbr_nd = wght_abund) # rename attribute of abundance 


# things to change in the following script:
# name of and text inside the maps

# ---------------------------- IMPORTANT-------------------------------------------
```

## Summary

This section aims to build a pipeline of spatial interpolation on feral pig abundance at TNC's Jack and Laura Dangermond Preserve (JLDP). Spatial interpolation uses data at certain points of observation to estimate the values of designated variable over the entire region of interest. In other words, it turns discrete point observations into a variable surface. For this project, we interpolate the number of feral pig abundance at `r length(unique(pig_abundance$statn_d))` camera stations operated during `r month_check` check (from `r first_day` to `r last_day`) to the entire JLDP. Analysis basically followed the [tutorial](https://mgimond.github.io/Spatial/interpolation-in-r.html) by Manuel Gimond. We sum up areas with most active pig visits, examine uncertainty of analysis, and give recommendations for further study of feral pig in the preserve. 
Note: Current version of analysis works with abundance data only so that it does not identify individual pigs.

## Analysis

### Feral pig abundance for each month

```{r}
# brief overview of the data
pig_sum <- pig_abundance %>% 
  group_by(statn_d) %>% 
  summarize(n_abundance = mean(nmbr_nd)) %>% 
  filter(n_abundance != "NA")
```


### Feral pig abundance of each camera station at JLDP

```{r}
station_coord <- as.data.frame(st_coordinates(pig_sum))
station_coord$statn_d = pig_sum$statn_d

abundance_map <- tm_shape(jldp_boundary) + 
  tm_polygons() +
  tm_shape(pig_sum) +
  tm_dots(col = "n_abundance",
          palette = "Reds", 
          auto.palette.mapping = FALSE,
          title = "Feral Pig Abundance \n(April check)",
          size = 0.3) + # tmaptools::palette_explorer() +
  tm_legend(legend.outside = TRUE)

abundance_map
```

``` {r }
# save the map
# change the figure name when needed
tmap::tmap_save(abundance_map, filename = here::here("figure", "abundance", "april", "abundance_map_april.png"))
```


### Thiessen polygons

```{r}
# convert sf shapefile to SpatialPointsDataFrame
jldp_boundary_sp <- sf::as_Spatial(jldp_boundary)

pig_sum_sp <- sf::as_Spatial(pig_sum)

# assign preserve boundary as window for wild pig abundance
pig_sum_sp@bbox <- jldp_boundary_sp@bbox

# create tessellated surface
pig_tess_surface <-as(dirichlet(as.ppp(pig_sum_sp)), "SpatialPolygons")
# mannually add projection information
sp::proj4string(pig_tess_surface) <- sp::proj4string(pig_sum_sp)

# add point attributes to tessellated surface
pig_tess_surface.z <- sp::over(pig_tess_surface, pig_sum_sp, fn = mean)
pig_tess_surface.spdf <- sp::SpatialPolygonsDataFrame(pig_tess_surface, pig_tess_surface.z)

# clip tessellated surface to JLDP boundary
pig_tess_surface.clp <- raster::intersect(jldp_boundary_sp, pig_tess_surface.spdf)

# map it
polygon_abundance_map <- tm_shape(pig_tess_surface.clp) +
  tm_polygons(col = "n_abundance",
              title = "Feral Pig Abundance \n(April check)") +
  tm_legend(legend.outside = TRUE)

polygon_abundance_map
```

``` {r }
# save the map
# change the figure name when needed
tmap::tmap_save(polygon_abundance_map, filename = here::here("figure", "abundance", "april", "polygon_abundance_map_april.png"))
```


### Inverse Distance Weighted (IDW)

```{r}
# create empty grid with n total cells in boundary box of pig_sum_sp
grd <- as.data.frame(sp::spsample(pig_sum_sp, "regular", n=50000))
names(grd) <- c("X", "Y")
sp::coordinates(grd) <- c("X", "Y")
sp::gridded(grd) <- TRUE # Create SpatialPixel object
sp::fullgrid(grd) <- TRUE # Create SpatialGrid object

# mannually add projection information
sp::proj4string(grd) <- sp::proj4string(pig_sum_sp)

# interpolate the grid cells using a power value of 2 (idp = 2)
pig_sum_idw <- gstat::idw(n_abundance ~ 1, pig_sum_sp, newdata = grd, idp = 2.0)

# convert to raster object 
pig_sum_idw_raster <- raster::raster(pig_sum_idw)

# clip to JLDP using mask
pig_sum_idw_clip <- raster::mask(pig_sum_idw_raster, jldp_boundary_sp)

# map it
idw_heatmap <- tm_shape(pig_sum_idw_clip) +
  tm_raster(n = 5, midpoint = 200,
            title = "Feral Pig abundances \n(April check)") +
  tm_shape(pig_sum) +
  tm_dots(size = 0.2) +
  tm_legend(legend.outside = TRUE)

idw_heatmap 
```

```{r}
# save the map
# change the figure name when needed
tmap::tmap_save(idw_heatmap, filename = here::here("figure", "abundance", "april", "idw_heatmap_april.png"))

writeRaster(pig_sum_idw_clip, filename = here::here("output_data", "abundance", "idw_heatmap_april.tif"), overwrite = TRUE)
```

```{r}
# leave-one-out validation
pig_idw_out <- vector(length = length(pig_sum_sp))

for (i in 1:length(pig_sum_sp)){
  pig_idw_out[i] <- gstat::idw(n_abundance ~ 1, pig_sum_sp[-i,], pig_sum_sp[i,], idp = 2.0)$var1.pred
}

# plot the differences
OP <- par(pty="s", mar=c(4,3,0,0))

plot(pig_idw_out ~ pig_sum_sp$n_abundance, 
     asp = 1,
     xlab = "Observed",
     ylab = "Predicted",
     pch = 16,
     col = rgb(0,0,0,0.5))
abline(lm(pig_idw_out ~ pig_sum_sp$n_abundance),
       col = "red",
       lw=2,
       lty=2)
abline(0,1)
par(OP)
```

```{r}
# compute RMSE
sqrt(sum((pig_idw_out - pig_sum_sp$n_abundance)^2)/length(pig_sum_sp))
```

### Cross validation

```{r}
# create the interpolated surface
img <- gstat::idw(n_abundance ~ 1, 
                  pig_sum_sp, 
                  newdata = grd, 
                  idp = 2.0)

n <- length(pig_sum_sp)
Z_i <- matrix(nrow = length(img$var1.pred),
              ncol = n)

# Remove a point, then interpolate
st <- stack()

for (i in 1:n) {
  Z_1 <- gstat::idw(n_abundance ~ 1,
                    pig_sum_sp[-i,],
                    newdata=grd,
                    idp=2.0)
  st <- addLayer(st, raster(Z_1, layer=1))
  Z_i[,i] <- n*img$var1.pred - (n-1)*Z_1$var1.pred # pseudo-value Z_j
}


# Jackknife estimator of parameter Z at location j
Z_j <- as.matrix(apply(Z_i, 1, sum, na.rm=TRUE)/n)


# Compute (Zi* - Zj)^2
c1 <- apply(Z_i,2,'-',Z_j)            # Compute the difference
c1 <- apply(c1^2, 1, sum, na.rm=T ) # Sum the square of the difference


# Compute the confidence interval
CI <- sqrt( 1/(n*(n-1)) * c1)

# Create (CI / interpolated value) raster
img.sig   <- img
img.sig$v <- CI /img$var1.pred 

# Clip the confidence raster to Texas
r <- raster(img.sig, layer="v")
r.m <- mask(r, jldp_boundary_sp)

ci95_map <- tm_shape(r.m) + 
  tm_raster(n=10,title="95% Confidence Interval for \nRange of Pig Abundance \n(April check)") +
  tm_shape(pig_sum_sp) + tm_dots(size=0.1) +
  tm_legend(legend.outside=TRUE)

ci95_map
```


``` {r }
# save the map
# change the figure name when needed
tmap::tmap_save(ci95_map, filename = here::here("figure", "abundance", "april", "ci95_map_april.png"))
```


### First order polynomial fit 

```{r}
# add coordinates to pig_sum_sp
pig_sum_sp$X <- coordinates(pig_sum_sp)[,1]
pig_sum_sp$Y <- coordinates(pig_sum_sp)[,2]

lm.1 <- lm(n_abundance ~ X + Y, data = pig_sum_sp)


# Use the regression model output
dat.1st <- sp::SpatialGridDataFrame(grd, data.frame(var1.pred = predict(lm.1, newdata = grd)))

# clip the interpolated raster

r_1st <- raster(dat.1st)
r_1st_m <- mask(r_1st, jldp_boundary_sp)

# Plot the map
tm_shape(r_1st_m) + 
  tm_raster(n=10, 
            title="Feral Pig Abundance (1st order) \n(April check)") +
  tm_shape(pig_sum_sp) + 
  tm_dots(size=0.2) +
  tm_legend(legend.outside=TRUE)

```

### 2nd order polynomial 

```{r}
lm.2 <- lm(n_abundance ~ X + Y + I(X*X)+I(Y*Y) + I(X*Y), data = pig_sum_sp)

dat.2nd <- SpatialGridDataFrame(grd, data.frame(var1.pred = predict(lm.2, newdata=grd))) 

# Clip the interpolated raster 
r_2nd   <- raster(dat.2nd)
r_2nd_m <- mask(r_2nd, jldp_boundary_sp)

# Plot the map
tm_shape(r_2nd_m) + 
  tm_raster(n=5, 
            title="Feral Pig Abundance (2nd order) \n(April check)") +
  tm_shape(pig_sum_sp) + 
  tm_dots(size=0.2) +
  tm_legend(legend.outside=TRUE)
```

### Kriging

#### Fit the variogram model

```{r}
# sample variogram 
var_detrend <- gstat::variogram(n_abundance ~ X + Y, pig_sum_sp, cloud = FALSE)

dat_fit  <- fit.variogram(var_detrend, fit.ranges = FALSE, fit.sills = FALSE,
                          vgm(psill=180, model="Sph", range=15000, nugget=0))

# The following plot allows us to assess the fit
plot(var_detrend, dat_fit, xlim=c(0,20000))
```

#### Kriging surface

```{r}
pig_krig <- krige(n_abundance ~ X + Y, pig_sum_sp, grd, dat_fit)

# clip the surface
r_krig_m <- mask(raster(pig_krig), jldp_boundary_sp)

krig_heatmap <- tm_shape(r_krig_m) +
  tm_raster(n = 5, 
            title = "Feral Pig Abundance - Kriging \n(April check)") +
  tm_shape(pig_sum_sp) +
  tm_dots(size = 0.2) +
  tm_legend(legend.outside = TRUE)

krig_heatmap
```

```{r}
# save the map
# change the figure name when needed
tmap::tmap_save(krig_heatmap, filename = here::here("figure", "abundance", "april", "krig_heatmap_april.png"))

writeRaster(r_krig_m, filename = here::here("output_data", "abundance", "krig_heatmap_april.tif"), overwrite = TRUE)
```


#### Confidence of analysis

##### Variance map
```{r}
r_var_m <- mask(raster(pig_krig, layer = "var1.var"), jldp_boundary_sp)

var_map <- tm_shape(r_var_m) +
  tm_raster(n = 7,
            palette ="Reds",
            title = "Variance Map \n(April check)") +
  tm_shape(pig_sum_sp) +
  tm_dots(size = 0.2) +
  tm_legend(legend.outside = TRUE)

var_map
```


``` {r }
# save the map
# change the figure name when needed
tmap::tmap_save(var_map, filename = here::here("figure", "abundance", "april", "var_map_april.png"))
```


##### Confidence interval map

```{r}
ci_pig <- sqrt(raster(pig_krig, layer = "var1.var")) * 1.96
ci_pig_m <- mask(ci_pig, jldp_boundary_sp)

ci95_krig_map <- tm_shape(ci_pig_m) +
  tm_raster(n = 7,
            palette = "Reds",
            title = "95% CI Map \n(April check)") +
  tm_shape(pig_sum_sp) +
  tm_dots(size = 0.2) +
  tm_legend(legend.outside = TRUE)

ci95_krig_map
```

``` {r }
# save the map
# change the figure name when needed
tmap::tmap_save(ci95_krig_map, filename = here::here("figure", "abundance", "april", "ci95_krig_map_april.png"))
```



## Conclusion and next steps

- Feral pigs are mostly active on the southeastern areas of JLDP during the summer season. 

- Survey for sounder size, sex ratio, average litter size may focus on areas with high pig abundance for better chance of sighting.

- Future camera stations may install around locations with larger CI and higher variance to improve understanding of pig activities. 

- Further analysis on monthly feral pig movement or comparation between winter and summer can be very helpful once the data is available. 


## Reference

- Tutorial of [spatial interpolation in R ](https://mgimond.github.io/Spatial/interpolation-in-r.html)
